The state of knowledge for marine invertebrate biodiversity in the continental US

### 2018-01-16 manuscript.Rmd
edit manuscript

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_us_inverts_records_species), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_us_inverts_records_species$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_us_inverts_records_species$worms_phylum))` phyla), and `r format(nrow(obis_us_inverts_records_species), big.mark=",")` records from OBIS (`r format(n_distinct(obis_us_inverts_records_species$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_us_inverts_records_species$worms_phylum))` phyla).

@@

Marine invertebrates are underrepresented in biodiversity databases. This underrepresentation is manifest in the small proportion of records they represent in the iDigBio database. While marine invertebrates compose between 11% and 18% of the described global diversity (number of described marine inverterbrates species: 200,000-210,000 [@Appeltans2012], total number of described species: 1.2-1.9 million [@Mora2011,@Chapman2009]),
marine invertebrates from the US EEZ only make up `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$total_no_fossil*100)`% of the total number of non-fossil iDigBio records (`r round(idigbio_stats$total_no_fossil/1e6, 1)` million), and `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$eez_no_fossil*100)`% of the US records (`r round(idigbio_stats$eez_no_fossil/1e6, 1)` million).

@@

Most samples have been collected on the continental shelf (Fig. \@ref(fig:map-diversity-sampling)A), with `r prop_idigbio_continental_shelf`% (iDigBio) and `r prop_obis_continental_shelf`% (OBIS) of the records collected between 0-200m. More records have been collected from the Atlantic Ocean, than the Gulf of Mexico, and the Pacific Ocean (`r format(sum(combined_records$is_east_coast), big.mark=",")`, `r format(sum(combined_records$is_within_gom), big.mark=",")`, and `r format(sum(combined_records$is_west_coast), big.mark=",")` records respectively). The highest species richness has been recorded from the Gulf of Mexico (`r format(n_spp_gom, big.mark=",")` species), followed by the Atlantic Ocean (`r format(n_spp_east, big.mark=",")` species), and the Pacific Ocean (`r format(n_spp_west, big.mark=",")` species).

@@

@@

```{r correlation-sampling, fig.cap='Relation between number of records and number of species for each cell of the map used in Fig. \\@ref(fig:map-diversity-sampling). The black line represents the 1:1 relationship. (\\#fig:correlation-sampling)'}
knitr::include_graphics("figures/plot_sampling_effort.pdf")
```

@@

Spatial patterns of species richness are however strongly confounded by sampling effort. There is a tight correlation between the number of samples and number of species per cell used in Fig. \@ref(fig:map-diversity-sampling), and `r format_output(stat_sampling_effort$p * 100)`% (n = `r stat_sampling_effort$n`) of the cells with $\leq$ 500 records have a number of species within 10% of the number of records (Fig. \@ref(fig:correlation-sampling)).

>>>>

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_us_inverts_records_species), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_us_inverts_records_species$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_us_inverts_records_species$worms_phylum))` phyla), and `r format(nrow(obis_us_inverts_records_species), big.mark=",")` occurence records from OBIS (`r format(n_distinct(obis_us_inverts_records_species$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_us_inverts_records_species$worms_phylum))` phyla).

@@

Given their diversity, marine invertebrates are underrepresented in iDigBio. While marine invertebrates are estimated to represent between 11% and 18% of the described global diversity (number of described marine inverterbrates species: 200,000-210,000 [@Appeltans2012], total number of described species: 1.2-1.9 million [@Mora2011,@Chapman2009]),
marine invertebrates from the US EEZ only make up `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$total_no_fossil*100)`% of the total number of non-fossil iDigBio records (`r round(idigbio_stats$total_no_fossil/1e6, 1)` million), and `r format_output(nrow(idigbio_us_inverts_records_species)/idigbio_stats$eez_no_fossil*100)`% of the US records (`r round(idigbio_stats$eez_no_fossil/1e6, 1)` million).

@@

More records have been collected from the Atlantic Ocean, than the Gulf of Mexico, and the Pacific Ocean (`r format(sum(combined_records$is_east_coast), big.mark=",")`, `r format(sum(combined_records$within_gom), big.mark=",")`, and `r format(sum(combined_records$is_west_coast), big.mark=",")` records respectively). The highest species richness has been recorded from the Gulf of Mexico (`r format(n_spp_gom, big.mark=",")` species), followed by the Atlantic Ocean (`r format(n_spp_east, big.mark=",")` species), and the Pacific Ocean (`r format(n_spp_west, big.mark=",")` species).

Most samples have been collected on the continental shelf (Fig. \@ref(fig:map-diversity-sampling)A), with `r prop_idigbio_continental_shelf`% (iDigBio) and `r prop_obis_continental_shelf`% (OBIS) of the records collected between 0-200m.

@@

Talk about contributions of different institutions here

@@

```{r correlation-sampling, fig.cap='Relationship between number of records and number of species for each cell of the map used in Fig. \\@ref(fig:map-diversity-sampling). The black line represents the 1:1 relationship. (\\#fig:correlation-sampling)'}
knitr::include_graphics("figures/plot_sampling_effort.pdf")
```

@@

Spatial patterns of species richness are however strongly confounded by sampling effort. There is a tight correlation between the number of samples and number of species per cell used in Fig. \@ref(fig:map-diversity-sampling), and `r format_output(stat_sampling_effort$p * 100)`% (n = `r format(stat_sampling_effort$n, big.mark=",")`) of the cells with $\leq$ 500 records have a number of species within 10% of the number of records (Fig. \@ref(fig:correlation-sampling)).
